---
title: "HW 03"
author: "Cedric Destin"
format:
  html:
    embed-resources: true
    execute:
      warning: false
      message: false
      error: false
toc: true
---

```{r setup}
#| label: Setup
options(warn=-1)
if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load(tidyverse,
               glue,
               scales,
               openintro,
               gridExtra,
               ggrepel,
               ggmap,
               ggridges,
               dsbox,
               devtools,
               fs,
               janitor,
               here,
               dplyr,
               palmerpenguins,
               stringr,
               ggplot2,
               plotly,
               Hmisc,
               ggExtra,
               patchwork,
               viridis,
               grid,
               gridExtra
               ) 
```

## 1 - Du Bois challenge.

```{r import_dubois_challenge}
#| label: import_dubois_challenge
income <- read_csv(here("data", "income.csv"), 
                    show_col_types = FALSE)

```

```{r clean_dubois_challenge}
#| label: clean_dubois_challenge
income <- income |> 
  mutate(
      Rent_pct = Rent / 100,
      Food_pct = Food / 100,
      Clothes_pct = Clothes / 100,
      Tax_pct = Tax / 100,
      Other_pct = Other / 100
    )
```

```{r add_average_income}
#| label: add_average_income
income <- income %>%
  mutate(Class = paste0(as.character(Class), "   $", as.character(Average_Income)))

```

```{r factor_class_order}
#| label: factor_class_order
income$Class <- factor(income$Class, levels = c("$1000 AND OVER   $1125", "$750-1000   $880", "$500-750   $547", "$400-500   $433.82", "$300-400   $335.66", "$200-300   $249.45", "$100-200   $139.1"))
```

Sources: https://stackoverflow.com/questions/18413756/re-ordering-factor-levels-in-data-frame

```{r}
glimpse(income)
```

```{r plot_income_expenditure}
#| label: plot_income_expenditure
# Your original code, modified to include text labels
fig <- plot_ly(income, x = ~Rent, y = ~Class, 
               type = 'bar', 
               orientation = 'h',
               name = 'Rent',
               text = ~Rent_pct,
               textposition = 'inside',
               texttemplate = '%{text:.0%}',
               marker = list(color = '#121210'))

fig <- fig %>% add_trace(x = ~Food, name = 'Food',
                         text = ~Food_pct,
                         textposition = 'inside',
                         texttemplate = '%{text:.0%}',
                         marker = list(color = '#7D5A7F'))

fig <- fig %>% add_trace(x = ~Clothes, name = 'Clothes',
                         text = ~Clothes_pct,
                         textposition = 'inside',
                         texttemplate = '%{text:.0%}',
                         marker = list(color = '#D79684'))

fig <- fig %>% add_trace(x = ~Tax, name = 'Tax',
                         text = ~Tax_pct,
                         textposition = 'inside',
                         texttemplate = '%{text:.1%}',
                         marker = list(color = '#003e80'))

fig <- fig %>% add_trace(x = ~Other, name = 'Other',
                         text = ~Other_pct,
                         textposition = 'inside',
                         texttemplate = '%{text:.1%}',
                         marker = list(color = '#e6f2ff'))

fig <- fig %>% layout(
  barmode = 'stack',
  title = "INCOME AND EXPENITURE OF 150 NEGRO FAMILIES IN ATLANTA, GA. U.S.A.",
  titlefont = list(size = 15, color = "#000000"),
  xaxis = list(
    title = "FOR FUTHER STATISTICS RAISE THIS FRAME",
    showticklabels = FALSE
  ),
    annotations = list(
      list(
        x = -0.28,
        y = 1.025,
        text = "Class     Actual Average",
        showarrow = FALSE,
        xref = "paper",
        yref = "paper"
      )
    ),
  yaxis = list(title = ""),
  showlegend = FALSE,
  plot_bgcolor = "#CAB2A0",
  paper_bgcolor = "#CAB2A0"
)

fig
```

Sources:

https://plotly.com/r/horizontal-bar-charts/ https://plotly.com/r/text-and-annotations

<https://stackoverflow.com/questions/66535691/plotly-how-to-automate-y-axis-with-space-for-texttemplate>

## 2 - COVID survey - interpret

The plot illustrates opinions made by several different groups regarding the COVID vaccine below are three observations I made: 1. Nurses seem to strongly recommend the vaccine, with a very small error bar, illustrating that most nurses are of the same mindset. In fact, most groups seem to recommend the vaccines except for a few groups, “Had COVID vaccine: No” and “Gender: Prefer not to say”. 2. Most groups were very confident in the scientific vetting the process for the new COVID vaccines, again, only a few groups showed disagreement to it, being the same groups disagreeing with the previous statement: “Had COVID vaccine: No” and “Gender: Prefer not to say”. 3. This time we could observe two very interesting groups, the “Had COVID vaccine” and the “Had flu vaccine this year” a. It is understandable that the people who had the COVID vaccine responded positively about the vaccine, like the people that had the flu vaccine this year, they mostly had positive feedback. b. On the other hand, the groups that did not have COVID vaccines had responses that were split straight in the middle, but with large amount of uncertainty. This illustrates that the consensus regarding the vaccine within that group is very broad, and that they may not have had the vaccines for a variety of reasons.

## 3 - COVID survey - reconstruct

```{r covid_survey_reconstruct}
#| label: COVID_survey_reconstruct
covid_survey <- read_csv(here("data", "covid-survey.csv"),
                         show_col_types = FALSE, skip = 1)
glimpse(covid_survey)

print(
  dim(covid_survey)
)

covid_survey <- covid_survey %>%
  filter(if_all(-response_id, ~ !is.na(.)))

print(
  dim(covid_survey)
)

```

```{r clean_covid_survey}
#| label: clean_covid_survey
covid_survey <- covid_survey %>%
  mutate(
    exp_already_vax = ifelse(exp_already_vax == 0, "No", "Yes"),
    exp_flu_vax = ifelse(exp_flu_vax == 0, "No", "Yes"),
    exp_profession = ifelse(exp_profession == 0, "Medical", "Nursing"),
    exp_gender = ifelse(exp_gender == 0, "Male", 
    ifelse(exp_gender == 1, "Female", 
            ifelse(exp_gender == 3, "Non-binary third gender", "Prefer not to say"))),
    exp_race = ifelse(exp_race == 1, "American Indian / Alaskan Native", 
                      ifelse(exp_race == 2, "Asian", 
                      ifelse(exp_race == 3, "Black or African American", 
                      ifelse(exp_race == 4, "Native Hawaiian / Other Pacific Islander", "White")))),
    exp_ethnicity = ifelse(exp_ethnicity == 1, "Hispanic / Latino", "Non-Hispanic/Non-Latino"),
    exp_age_bin = case_when(
      exp_age_bin == 0 ~ "<20",
      exp_age_bin == 20 ~ "21-25",
      exp_age_bin == 25 ~ "26-30",
      exp_age_bin == 30 ~ ">30"
    )
  )

print(
  dim(covid_survey)
)
```

```{r pivot_covid_survey}
#| label: pivot_covid_survey
covid_survey_longer <- covid_survey |>
  pivot_longer(
    cols = starts_with("exp_"),
    names_to = "explanatory",
    values_to = "explanatory_value"
  ) |>
  filter(!is.na(explanatory_value)) |>
  pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
  )

print(covid_survey_longer)
```

```{r group_covid_survey}
#| label: group_covid_survey
covid_survey_summary_stats_by_group <- covid_survey_longer %>%
  group_by(explanatory, explanatory_value, response) %>%
  summarise(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), probs = 0.10, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.90, na.rm = TRUE)
  )

print(covid_survey_summary_stats_by_group)
```

```{r summarise_covid_survey}
#| label: summarise_covid_survey
covid_survey_summary_stats_all <- covid_survey_longer %>%
  group_by(response) %>%
  summarise(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), probs = 0.10, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.90, na.rm = TRUE),
    explanatory = "All",
    explanatory_value = ""
  )

print(covid_survey_summary_stats_all)
```

```{r bind_dataframes}
#| label: bind_dataframes
covid_survey_summary_stats <- bind_rows(
  covid_survey_summary_stats_by_group,
  covid_survey_summary_stats_all
)

print(covid_survey_summary_stats)
```

```{r mutate_covid_survey}
#| label: mutate_covid_survey
covid_survey_summary_stats  <- covid_survey_summary_stats %>%
  mutate(
    explanatory = case_when(
      explanatory == "exp_age_bin" ~ "Age",
      explanatory == "exp_already_vax" ~ "Had COVID vaccine",
      explanatory == "exp_flu_vax" ~ "Had flu vaccine this year",
      explanatory == "exp_profession" ~ "Profession",
      explanatory == "exp_gender" ~ "Gender",
      explanatory == "exp_race" ~ "Race",
      explanatory == "exp_ethnicity" ~ "Ethnicity",
      explanatory == "All" ~ "All"
      ),
      response = case_when(
      response == "resp_safety" ~ "Based on my understanding, I believe the vaccine is safe",
      response == "resp_confidence_science" ~ "I am confident in the scientific vetting process for the new COVID vaccines",
      response == "resp_feel_safe_at_work" ~ "Getting the vaccine will make me feel safer at work",
      response == "resp_will_recommend" ~ "I will recommend the vaccine to family, friends, and community members",
      response == "resp_trust_info" ~ "I trust the information that I have received about the vaccines",
      response == "resp_concern_safety" ~ "I am concerned about the safety and side effects of the vaccine"
    )
  )
print(covid_survey_summary_stats)
```

```{r change_order}
#| label: change_order
custom_order <- c("All", "Age", "Gender", "Race", "Ethnicity", "Profession", "Had COVID vaccine", "Had flu vaccine this year")
covid_survey_summary_stats$explanatory <- factor(covid_survey_summary_stats$explanatory, levels = custom_order)
covid_survey_summary_stats_sorted <- covid_survey_summary_stats[order(covid_survey_summary_stats$explanatory), ]
print(covid_survey_summary_stats_sorted)
```

```{r change_order_level_2}
#| label: change_order_level_2
custom_order <- c("", ">30", "26-30", "21-25", "<20", "Prefer not to say", "Non-binary third gender", "Male", "Female", 
"White", "Native Hawaiian / Other Pacific Islander", "Black or African American", "Asian",  "American Indian / Alaskan Native", 
"Non-Hispanic/Non-Latino", "Hispanic / Latino", "Nursing", "Medical", "Yes", "No")
covid_survey_summary_stats_sorted$explanatory_value <- factor(covid_survey_summary_stats_sorted$explanatory_value, levels = custom_order)
covid_survey_summary_stats_sorted <- covid_survey_summary_stats_sorted[order(covid_survey_summary_stats_sorted$explanatory_value), ]
print(covid_survey_summary_stats_sorted)
```

```{r plot_covid_survey, fig.width=10, fig.height=7}
#| label: plot_covid_survey
ggplot(covid_survey_summary_stats_sorted, aes(x = mean, y = explanatory_value)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.2, position = position_dodge(width = 0.5)) +
  facet_grid(cols = vars(response), 
              rows = vars(explanatory),
              labeller = labeller(response = label_wrap_gen(15),
                                  explanatory = label_wrap_gen(15)),
              space = "free_y",
              scales = "free_y") +
  labs(x = "Mean likert score \n (Error bars range from 10th to 90th percentile)", y = "") + 
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "gray90", color = "lightgray"),
    strip.text.x = element_text(angle = 0),
    strip.text.y = element_text(angle = 0)
  ) + 
  removeGrid()
```

Overall, I do not see a great difference between the 10 - 90th percentile plot and the 25 - 75th percentile plot. the error bar on the 25 - 75th percentile plot is narrower, but I do not think it is adding much value to the plot that the 10 - 90th percentile plot already provides. Hence my conclusion still stands that the overall sentiment towards the COVID vaccine is positive.

Sources:

<https://www.datamentor.io/r-programming/ifelse-function>

<https://www.geeksforgeeks.org/switch-case-in-r/>

<https://ggplot2.tidyverse.org/reference/labellers.html>

<https://ggplot2.tidyverse.org/reference/geom_errorbarh.html>

## 4 - COVID survey - re-reconstruct

```{r covid_survey_re_reconstruct}
#| label: covid_survey_re_reconstruct
covid_survey_summary_stats_all <- covid_survey_longer %>%
  group_by(response) %>%
  summarise(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), probs = 0.25, na.rm = TRUE),
    high = quantile(as.numeric(response_value), probs = 0.75, na.rm = TRUE),
    explanatory = "All",
    explanatory_value = ""
  )

print(covid_survey_summary_stats_all)
```

```{r bind_dataframes_2}
#| label: bind_dataframes_2
covid_survey_summary_stats <- bind_rows(
  covid_survey_summary_stats_by_group,
  covid_survey_summary_stats_all
)

print(covid_survey_summary_stats)
```

```{r mutate_covid_survey_2}
#| label: mutate_covid_survey_2
covid_survey_summary_stats  <- covid_survey_summary_stats %>%
  mutate(
    explanatory = case_when(
      explanatory == "exp_age_bin" ~ "Age",
      explanatory == "exp_already_vax" ~ "Had COVID vaccine",
      explanatory == "exp_flu_vax" ~ "Had flu vaccine this year",
      explanatory == "exp_profession" ~ "Profession",
      explanatory == "exp_gender" ~ "Gender",
      explanatory == "exp_race" ~ "Race",
      explanatory == "exp_ethnicity" ~ "Ethnicity",
      explanatory == "All" ~ "All"
      ),
      response = case_when(
      response == "resp_safety" ~ "Based on my understanding, I believe the vaccine is safe",
      response == "resp_confidence_science" ~ "I am confident in the scientific vetting process for the new COVID vaccines",
      response == "resp_feel_safe_at_work" ~ "Getting the vaccine will make me feel safer at work",
      response == "resp_will_recommend" ~ "I will recommend the vaccine to family, friends, and community members",
      response == "resp_trust_info" ~ "I trust the information that I have received about the vaccines",
      response == "resp_concern_safety" ~ "I am concerned about the safety and side effects of the vaccine"
    )
  )
print(covid_survey_summary_stats)
```

```{r change_order_2}
#| label: change_order_2
custom_order <- c("All", "Age", "Gender", "Race", "Ethnicity", "Profession", "Had COVID vaccine", "Had flu vaccine this year")
covid_survey_summary_stats$explanatory <- factor(covid_survey_summary_stats$explanatory, levels = custom_order)
covid_survey_summary_stats_sorted_2 <- covid_survey_summary_stats[order(covid_survey_summary_stats$explanatory), ]
print(covid_survey_summary_stats_sorted_2)
```

```{r change_order_level_2_2}
#| label: change_order_level_2_2
custom_order_2 <- c("", ">30", "26-30", "21-25", "<20", "Prefer not to say", "Non-binary third gender", "Male", "Female", 
"White", "Native Hawaiian / Other Pacific Islander", "Black or African American", "Asian",  "American Indian / Alaskan Native", 
"Non-Hispanic/Non-Latino", "Hispanic / Latino", "Nursing", "Medical", "Yes", "No")
covid_survey_summary_stats_sorted_2$explanatory_value <- factor(covid_survey_summary_stats_sorted_2$explanatory_value, levels = custom_order_2)
covid_survey_summary_stats_sorted_2 <- covid_survey_summary_stats_sorted_2[order(covid_survey_summary_stats_sorted_2$explanatory_value), ]
print(covid_survey_summary_stats_sorted_2)
```

```{r plot_covid_survey_2, fig.width=10, fig.height=7}
#| label: plot_covid_survey_2
ggplot(covid_survey_summary_stats_sorted_2, aes(x = mean, y = explanatory_value)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.2, position = position_dodge(width = 0.5)) +
  facet_grid(cols = vars(response), 
              rows = vars(explanatory),
              labeller = labeller(response = label_wrap_gen(15),
                                  explanatory = label_wrap_gen(15)),
              space = "free_y",
              scales = "free_y") +
  labs(x = "Mean likert score \n (Error bars range from 25th to 75th percentile)", y = "") + 
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "gray90", color = "lightgray"),
    strip.text.x = element_text(angle = 0),
    strip.text.y = element_text(angle = 0)
  ) + 
  removeGrid()
  
```

Sources:

<https://stackoverflow.com/questions/68989228/sort-dataframe-by-column-value-r>

## 5 - COVID survey - another view

```{r covid_survey_fixing_names}
#| label: covid_survey_fixing_names
covid_survey_longer <- covid_survey_longer %>%
  mutate(
    response_value = as.numeric(response_value),
    explanatory = case_when(
      explanatory == "exp_age_bin" ~ "Age",
      explanatory == "exp_already_vax" ~ "Had COVID vaccine",
      explanatory == "exp_flu_vax" ~ "Had flu vaccine this year",
      explanatory == "exp_profession" ~ "Profession",
      explanatory == "exp_gender" ~ "Gender",
      explanatory == "exp_race" ~ "Race",
      explanatory == "exp_ethnicity" ~ "Ethnicity"
    ),
    response = case_when(
      response == "resp_safety" ~ "Based on my understanding, I believe the vaccine is safe",
      response == "resp_confidence_science" ~ "I am confident in the scientific vetting process for the new COVID vaccines",
      response == "resp_feel_safe_at_work" ~ "Getting the vaccine will make me feel safer at work",
      response == "resp_will_recommend" ~ "I will recommend the vaccine to family, friends, and community members",
      response == "resp_trust_info" ~ "I trust the information that I have received about the vaccines",
      response == "resp_concern_safety" ~ "I am concerned about the safety and side effects of the vaccine"
    )
  ) 
print(covid_survey_longer)
```

```{r}
covid_survey_longer_grouped <- covid_survey_longer %>%
  group_by(response, response_value) %>%
  summarise(total = sum(response_value), .groups = "drop") %>%
  group_by(response) %>%
  mutate(pct = total / sum(total)) %>%
  ungroup()
```

```{r change_number_tostrings}
#| label: change_number_tostrings
covid_survey_longer_grouped <- covid_survey_longer_grouped %>%
mutate(
    response_value = as.character(response_value),
    response_value = case_when(
      response_value == "1" ~ "Strongly Agree",
      response_value == "2" ~ "Somewhat Agree",
      response_value == "3" ~ "Neither Agree Nor Disagree",
      response_value == "4" ~ "Somewhat Disagree",
      response_value == "5" ~ "Strongly Disagree"
    )
  )
```

```{r plot_covid_survey_all_stacked, fig.width=28, fig.height=5}
#| label: plot_covid_survey_all_stacked
custom_order_4 <- c("Strongly Agree", "Somewhat Agree", "Neither Agree Nor Disagree", "Somewhat Disagree", "Strongly Disagree")
covid_survey_longer_grouped$response_value <- factor(covid_survey_longer_grouped$response_value, levels = custom_order_4)
covid_survey_longer_grouped <- covid_survey_longer_grouped[order(covid_survey_longer_grouped$response_value), ]

p1 <- ggplot(covid_survey_longer_grouped, aes(y = pct, fill = response_value, x = "")) +
  coord_flip() +
  scale_fill_manual(values = c("Strongly Agree" = "#00a40f", 
                              "Somewhat Agree" = "#58d68d", 
                              "Neither Agree Nor Disagree" = "#fbf700", 
                              "Somewhat Disagree" = "#fb8100", 
                              "Strongly Disagree" = "#fe001b")) +
  geom_bar(position="stack", stat="identity") +
  labs(
       x = "", 
       y = "Proportion of Responses", 
      fill = "") +
  facet_grid(rows = vars(response), 
              labeller = labeller(response = label_wrap_gen(15),
                                  explanatory = label_wrap_gen(15)),
              space = "free_y",
              scales = "free_y") + 
  theme(
    strip.background = element_rect(fill = "gray90", color = "lightgray"),
    strip.text.y = element_text(angle = 0)
  ) +
  scale_y_continuous(labels = scales::percent) + 
  theme(
  strip.text.x = element_blank(),
  strip.text.y = element_blank()
) 
```

```{r}
covid_survey_longer_grouped <- covid_survey_longer_grouped %>%
  mutate(
    pct = case_when(
      response_value == "Strongly Disagree" ~ pct*-1,
      response_value == "Somewhat Disagree" ~ pct*-1,
      response_value == "Neither Agree Nor Disagree" ~ pct*-1,
      response_value == "Somewhat Agree" ~ pct,
      response_value == "Strongly Agree" ~ pct,
    )
  )
```

```{r fig.width=28, fig.height=5}
custom_order_4 <- c("Strongly Disagree", "Somewhat Disagree", "Neither Agree Nor Disagree", "Strongly Agree", "Somewhat Agree")
covid_survey_longer_grouped$response_value <- factor(covid_survey_longer_grouped$response_value, levels = custom_order_4)
covid_survey_longer_grouped <- covid_survey_longer_grouped[order(covid_survey_longer_grouped$response_value), ]


p2 <- ggplot(covid_survey_longer_grouped,aes(x=pct,
              fill=response_value,y=response))+
    geom_bar(stat='identity',width=0.5) +
    scale_y_discrete(labels = label_wrap(width = 25)) +
    scale_x_continuous(labels = scales::percent) + 
    scale_fill_manual(values = c("Strongly Agree" = "#00a40f", 
                              "Somewhat Agree" = "#58d68d", 
                              "Neither Agree Nor Disagree" = "#fbf700", 
                              "Somewhat Disagree" = "#fb8100", 
                              "Strongly Disagree" = "#fe001b")) +
    theme(legend.position = "none") +
    labs(
       x = "Proportion of Responses", y = " ") 

```

```{r}
custom_title <- textGrob(
  "Sentiments regarding the COVID-19 Vaccines",
  gp = gpar(
    fontfamily = "serif",
    fontsize = 22, 
    fontface = "bold", 
    col = "darkblue" 
  )
)
```

```{r final_plot, fig.width=28, fig.height=10}
grid.arrange(p2, p1, ncol = 2, top = "Sentiments regarding the COVID-19 Vaccines")                            
```

Both the diverging bar chart and the stacked bar chart provide a clear understanding of the sentiment towards the COVID vaccine. The diverging bar plot shows that the "Strongly Agree" responses dominate on the plot. Although the stack bar chart illustrates the same information, it also shows that in certain cases, the sentiment is neutral overall. In other words, every answers are equally distributed, which a bit harder to see in the diverging chart.

Sources:

<https://rdrr.io/cran/ggpubr/man/text_grob.html>